title: "Getting and Cleaning Data Course Project"
author: "Kevin Roy"
date: "01/07/2020"
output: html_document

# Explanation of the data set can be found in the "README.txt".

# Folder containing the data set from - "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
 
# I moved the UCI HAR folder to my working directory and out of the zip file

# I created the following data frames from this file. x contains the data, y contains the activity performed (numbered 1 to 6) and z contains the subject number who performed the activity (numbered 1 to 30)

# I combined these as the original data set was split into two data sets, with 70% of the data in the training dataset and 30% in the test dataset.In order to not mix up the data, I combined the training data set above the test data set.
```{r}
setwd("~/R/R Directory/John Hopkins/UCI HAR Dataset/test")

xtest <- read.delim("x_test.txt", sep = "", header =FALSE)
ytest <- read.delim("y_test.txt", sep = "", header = F)
subjecttest <- read.delim("subject_test.txt", header = F)
```

```{r}
setwd("~/R/R Directory/John Hopkins/UCI HAR Dataset/train")
xtrain <- read.delim("x_train.txt", sep = "", header = FALSE)
ytrain <- read.delim("y_train.txt", sep = "", header = FALSE)
subjecttrain <- read.delim("subject_train.txt", header = FALSE)

```
# I combined the tasks data frame by row. 
# I followed this up by changing the numbers of the task to the verb describing the task being measured. I returned it back to a data frame after converting it to a vector to perform the function. This way it will be easier to read, instead of a number format.
```{r}
tasks <- rbind(ytrain, ytest)

tasks <- as.vector(tasks$V1)
tasks <- mapvalues(tasks, from = 1:6, to = c("WALKING","WALKING_UPSTAIRS", "WALKING_DOWNSTAIRS", "SITTING", "STANDING", "LAYING"))

tasks <- as.data.frame(tasks)
```
# I combined the subjects from the two data sets and renamed the header to "subject". This includes the 30 adults between 18 and 48 years old who took part in the experiment.
```{r}
all_subjects <- rbind(subjecttrain, subjecttest)

all_subjects <- rename(all_subjects, subject = V1)

```

# I read in the following text document as it contains all the  header for a data set of "x_train.txt" and "x_test.txt". These include the mean, median, standard deviation, etc for each task completed.
```{r}
setwd("~/R/R Directory/John Hopkins/UCI HAR Dataset")

variables <- read.delim("features.txt", header = FALSE)
```
# I split the character vector into 2 parts. I kept the second element of it's name and removed the first element, as this contained the column number and was not necessary.
```{r}
splitNames <- strsplit(variables$V1," " )
splitNames[[1]][2]

secondElement <- function(x){x[2]}
variables <- sapply(splitNames,secondElement)
```
# The above function change the data frame into a character vector, so I returned it to a data frame
```{r}
variables <- as.data.frame(variables)

```
# I renamed the headers and changed them so that they were easier to read. hose columns that started with t was changed to time and f for frequency, also acceleration, magnitude and the gyroscope was written in full.
```{r}
variables$variables <- sub("^t", "time", variables$variables)
variables$variables <- sub("^f", "frequency", variables$variables)
variables$variables <- sub("[Aa]cc", "acceleration", variables$variables)
variables$variables <- sub("[Mm]ag", "magnitude", variables$variables)
variables$variables <- sub("[Gg]yro", "Gyroscope", variables$variables)
```

# The variable headers can be explained as follows:
# time in 50hz
# frequency in 20hz
# acceleration, which is recorded via an accelerometer
# mag - magnetism calculated using Euclidean norm
# gyroscope - orientation and angular velocity in 50hz
# x,y,z represents movement along x, y and z axis
# For more details on variables, please read the "features info" in data set



# I combined the x data sets together, which contains all the data values and attached the column names from the variables data set.
```{r}
totalresults <- rbind(xtrain,xtest)
colnames(totalresults) <- variables$variables
```
# I only want the data containing the mean and the standard deviation. I therefore subsetted the columns that had the name "mean" and "std" in the header
```{r}
totalresults<- select(totalresults, grep("[Mm][Ee][Aa][Nn]|[Ss][Tt][Dd]", names(totalresults), value = TRUE))
```


# I then combined the three data frames I have created, which included the data of time/frequency, the subject and the task completed.
```{r}
humanactivityrecognition <- cbind(all_subjects, tasks, totalresults)

```
# Lastly I took the average of each column based on the subject and task, then ordered them by task and subject.
```{r}
averagehumanactivityrecognition <- humanactivityrecognition %>% 
      group_by(subject, tasks) %>% 
      summarise_all(list(mean)) %>% 
      arrange(tasks, subject)
```